### 2009-11-15 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Updating the previous chapters 4, 5, and adding 6 and 7 unfinished.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@645 559edde2-a4d4-44e0-96fc-4035e40d7396

>>>>

\lstset{label=file:mongodb-ysi-data-format,caption=JSON representation of the
data produced by a YSI sensor on mongoDB}
\begin{lstlisting}
            {
             "_id" :  ObjectId( "d36f4007b7e7ac4a03c60000")  , 
             "sensor_ip_address" : "192.168.0.136", 
             "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b",
             "transaction_time" : 1252845473412 , 
             "fact_time" : 1252845346000,
             "latitude" : 12.3450, 
             "longitude" : -123.4456 , 
             "data" : {
                       "temperature" : 45.01, 
                       "sp_condition" : 37.6, 
                       "condition" : 145.8, 
                       "resistence" : 159.77, 
                       "salinitude" : 0.0, 
                       "pressure" : 0.391, 
                       "depth" : 0.46, 
                       "ph" : 5.64, 
                       "pH_mv" : -62.1, 
                       "odo_sat" : 89.7, 
                       "odo_condition" : 59.34, 
                       "turbidity" : 0.0,
                       "battery" : 9.4
                      }
            }
\end{lstlisting}

\lstset{label=file:rtc-ysi-opendap,caption=HTTP GET Request x Response to the OPeNDAP server at SF-BEAMS}
\begin{lstlisting}
http://sfbeams.sfsu.edu:8080/opendap/sfbeams/data_ctd/rtc_ctd5-ysimoor/real-time/sfb_CTD5_PUF.dat.ascii?
GET /opendap/sfbeams/data_ctd/rtc_ctd5-ysimoor/real-time/sfb_CTD5_PUF.dat.dods?&Month=11&Day=12 HTTP/1.1

HTTP/1.x 200 OK
Server: Apache-Coyote/1.1
Last-Modified: Thu, 12 Nov 2009 17:55:06 GMT
XDODS-Server: Server-Version-Unknown
XOPeNDAP-Server: bes/3.5.1 freeform_handler/3.7.5, netcdf_handler/3.7.6
XDAP: 3.2
Content-Description: dods_data
Content-Type: text/txt
Date: Sun, 15 Nov 2009 03:08:45 GMT

Dataset: sfb_CTD5_PUF.dat
YSI_REALTIME_CSV.Month, YSI_REALTIME_CSV.Day, YSI_REALTIME_CSV.Year, YSI_REALTIME_CSV.Hour, YSI_REALTIME_CSV.Min, YSI_REALTIME_CSV.Sec, YSI_REALTIME_CSV.WaterTemperature, YSI_REALTIME_CSV.SpecificConductivity, YSI_REALTIME_CSV.Conductivity, YSI_REALTIME_CSV.Resistivity, YSI_REALTIME_CSV.TDS, YSI_REALTIME_CSV.Salinity, YSI_REALTIME_CSV.Pressure, YSI_REALTIME_CSV.Depth, YSI_REALTIME_CSV.pH, YSI_REALTIME_CSV.pHmV, YSI_REALTIME_CSV.Turbidity, YSI_REALTIME_CSV.ODOSaturation, YSI_REALTIME_CSV.ODO, YSI_REALTIME_CSV.Chlor, YSI_REALTIME_CSV.Chlor_RFU, YSI_REALTIME_CSV.Battery, YSI_REALTIME_CSV.InstSN
11, 12, 2009, 16, 30, 51, 13.78, 4.3945, 3.4532, 0, 28.56, 28.35, 2.65, 2.651, 7.9, -54.7, 9, -99999, -99999, 3.7, 1, 9.4, -99999
11, 12, 2009, 16, 36, 50, 13.78, 4.4095, 3.4642, 0, 28.66, 28.46, 2.628, 2.63, 7.91, -54.9, 11.9, -99999, -99999, 4, 1.1, 9.4, -99999
11, 12, 2009, 16, 42, 51, 13.77, 4.3943, 3.4515, 0, 28.56, 28.35, 2.631, 2.632, 7.9, -54.9, 9.8, -99999, -99999, 3.6, 1, 9.4, -99999
11, 12, 2009, 16, 48, 50, 13.76, 4.3893, 3.447, 0, 28.53, 28.32, 2.628, 2.63, 7.9, -54.8, 10.6, -99999, -99999, 3.8, 1, 9.4, -99999
\end{lstlisting}

### 2009-11-15 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Updating the previous chapters 4, 5, and adding 6 and 7 unfinished.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@645 559edde2-a4d4-44e0-96fc-4035e40d7396

As it is shown, the data produced by this type of sensor runs in the order of
at most hundreds of Megabytes a year, what can be considered a very low data
storage requirement as compared to other systems such as web applications. Upon
collecting the data from sensors, the RTC lab staff use automation scripts
written in Matlab to process, index and distribute the raw data in different
formats, including the OPeNDAP \cite{opendap}, a format widely used in the
research institutions that promotes easier data exchange among them. The
access to the archived data is done by using the OPeNDAP server is located at
http://sfbeams.sfsu.edu:8080/opendap.

@@

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.6]{../diagrams/DSP-Property-Class-Diagram}
  \caption{The DSP Property Payload - Message Content UML Class Diagram}
  \label{fig:DSP-Property-Class-Diagram}
\end{figure}

@@

@@

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-Message-Broker-Abstraction}
  \caption{A DSP Message Broker Abstraction}
  \label{fig:DSP-Message-Broker-Abstraction}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSPBroker-Matcher-Class-Diagram}
  \caption{The DSP Broker-Matcher UML Class Diagram}
  \label{fig:DSPBroker-Matcher-Class-Diagram}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.6]{../diagrams/DSP-to-DSP-Remote-Communication}
  \caption{The remote communication between two DSP Components using the HTTP
  Protocol}
  \label{fig:DSP-to-DSP-Remote-Communication}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}
  \caption{The UML Sequence Diagram for the production of a YSI Message}
  \label{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}
\end{figure}

>>>>

As it is shown, the data produced by this type of sensor runs in the order of
at most hundreds of Megabytes a year, what can be considered a very low data
storage requirement as compared to other systems such as web applications. Upon
collecting the data from sensors, the RTC lab staff use automation scripts
written in Matlab to process, index and distribute the raw data in different
formats, including the OPeNDAP \cite{opendap}, a format widely used in the
research institutions that promotes easier data exchange among them. The
access to the archived data is done by using the OPeNDAP server is located at
http://sfbeams.sfsu.edu:8080/opendap, as a request and the related output for
one of the YSI devices can be see in listing \ref{file:rtc-ysi-opendap}.

@@

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.6]{../diagrams/DSP-Property-Class-Diagram}
  \caption{The DSP Property Payload - Message Content UML Class Diagram}
  \label{fig:DSP-Property-Class-Diagram}
\end{figure}

@@

\newpage

@@

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.8]{../diagrams/DSP-Message-Broker-Abstraction}
  \caption{A DSP Message Broker Abstraction}
  \label{fig:DSP-Message-Broker-Abstraction}
\end{figure}

@@

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.6]{../diagrams/DSPBroker-Matcher-Class-Diagram}
  \caption{The DSP Broker-Matcher UML Class Diagram}
  \label{fig:DSPBroker-Matcher-Class-Diagram}
\end{figure}

@@

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.6]{../diagrams/DSP-to-DSP-Remote-Communication}
  \caption{The remote communication between two DSP Components using the HTTP
  Protocol}
  \label{fig:DSP-to-DSP-Remote-Communication}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}
  \caption{The UML Sequence Diagram for the production of a YSI Message}
  \label{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}
\end{figure}

### 2009-11-15 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
Updating the previous chapters 4, 5, and adding 6 and 7 unfinished.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@645 559edde2-a4d4-44e0-96fc-4035e40d7396

Other projects have been using the XML data models. It falls into the same
category as the Relational Model, since the XML documents, when in the context
of a database, must be complaint to an XML Schema. As shown in section 3, this
model uses XPath technology for querying documents, although some hibrid
technologies may still use SQL for that matter.

@@

The advance of Cloud Computing, and dynamic web applications have motivated the
database community to implement different mechanisms for persisting data
without the heavy-process of data modeling. Given the fact that sensor networks
may be updated at any time by the addition of new sensor types, the use of
moding approaches such as Key-Value Pairs have been proposed []. One could be
tempted to implement a KVP over relational \ref{fig:KVP-on-Relational-Model}:

\begin{figure}
  \centering
  \includegraphics[scale=0.65]{../diagrams/KVP-on-Relational-Model}
  \caption{KVP Data Model implementation using Relational Model}
  \label{fig:KVP-on-Relational-Model}
\end{figure}

\begin{itemize}
  \item tules are not complaint to the normalization rules;
  \item difficult access to key and values
\end{itemize}

The real use of KVP and Document-Oriented data models mitigates the problem of
data modeling process.

@@

@@

>>>>

Other projects have been using the XML data models. It falls into the same
category as the Relational Model, since the XML documents, when in the context
of a database, must be complaint to an XML Schema. As shown in section 3, this
model uses XPath technology for querying documents, although some hibrid
technologies may still use SQL for that matter. In this way, an instance of
such schema dependent types can be seen in figure
\ref{fig:persistence-example-relational}.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.65]{../diagrams/persistence-example-relational}
  \caption{Instance of a Relational Model Database Prototype}
  \label{fig:persistence-example-relational}
\end{figure}

Although this data model have been used in different types of applications,
developers have tried to use the Relational Model to model an application that
could prevent any database change through the use of a technique that relates a
key to a value, called Key-Value data model. With the advent of Internet
applications, the need of such a model that does not need constant refactoring
led developers to propose such approach using the relational model as
documented in technical articles such as \cite{db-kvp-in-relational01} and
\cite{db-kvp-in-relational02}. Based in these articles, consider the relational
model depicted in figure \ref{fig:KVP-on-Relational-Model} as model for the
persistence data for our case study using the key-value strategy. The problem
with it is that key repetition occurs, since the nature of time-series data
requires the use of a timestamp key for each property of the sensor. Therefore,
this strategy is not well-suited to provide persistence data from collected
sensor networks.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/KVP-on-Relational-Model}
  \caption{KVP Data Model implementation using Relational Model}
  \label{fig:KVP-on-Relational-Model}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/persistence-example-relational-kvp}
  \caption{KVP over Relational Model Database Prototype}
  \label{fig:persistence-example-relational-kvp}
\end{figure}

@@

The advance of Cloud Computing, and dynamic web applications have motivated the
database community to implement different mechanisms for persisting data
without the heavy-process of data modeling. Given the fact that sensor networks
may be updated at any time by the addition of new sensor types, the use of
moding approaches such as Key-Value Database have been proposed \cite{db-kvp}.

\begin{itemize}
  \item entities from the same domains are placed into a bucket;
  \item entities have a set of attributes and relating values
  \item entities with different set of attributes may be contained in the same
  bucket, since there is no schema to govern the bucket items restriction.
\end{itemize}

For instance, all data needed for the YSI sonde data, as well as all necessary
provenance data that describes the data, are represented by means of key-value
pairs \ref{fig:persistence-example-kvp}.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/persistence-example-kvp}
  \caption{KVP Database instance Prototype}
  \label{fig:persistence-example-kvp}
\end{figure}

@@

The data models were compared by \cite{db-is-rdbs-dommed} and can be summarized
as follows:

\begin{table}
    \label{tab:ysi-data-distribution}
    \caption{Schema-Dependent X Schema-less Databases Compared: Properties}
    \begin{center}
    \begin{tabular}{|p{210pt}|p{210pt}|}\hline
    Schema-Dependent Databases & Schema-less Databases\\\hline
    \begin{enumerate}
      \item Real-world model by entities, classified in tables;
      \item Tables composed by columns and rows. Rows are comprised by column
      values, which have the same schema;
      \item Data Model defined in advance with a schema, which contains
      relationships and constraints to enforce data integrity;
      \item Data represents ``natural'' entities, not application specific;
      \item Data Normalization is the data structuring process to remove data
      duplication, as well as establishing data associations through table
      relationships.
    \end{enumerate} 
    & 
    \begin{enumerate}
      \item Real-world model by entities, classified in Domains;
      \item Domains are similar to tables, but like buckets that contains items
      without a pre-defined schema, what enable them to have different schemas;
      \item Items are identified by keys, as well as have a dynamic set of
      attributes attached to it, however with no schema defined;
      \item Items may represent not only the natural representation of data, but
      also include application-specific data;
      \item Since data may be repeated, no data normalization is done, so that
      integrity is done in the application layer.
    \end{enumerate}
    \\\hline
    \end{tabular}
    \end{center}
\end{table}

\begin{table}
    \label{tab:ysi-data-distribution}
    \caption{Schema-Dependent X Schema-less Databases: Data Access}
    \begin{center}
    \begin{tabular}{|p{210pt}|p{210pt}|}\hline
    Schema-Dependent Databases & Schema-less Databases\\\hline
    \begin{enumerate}
      \item The basic operations CRUD \footnote{Create-Retrieve-Update-Update
      database operations} data are performed using a structured language such
      as the SQL or XPath;
      \item Query languages can access data from different tables through
      joins, contains functions for aggregation and complex filter.
    \end{enumerate} 
    & 
    \begin{enumerate}
      \item The CRUD operations are performed via API \footnote{Application
      Programming Interface} through programming languages;
      \item Some technologies provide basic SQL-like syntax for filter criteria
      with some predicates like =, !=, <, > that ca be applied;
      \item The data and application integrity logic is placed in the
      application layer.
    \end{enumerate}
    \\\hline
    \end{tabular}
    \end{center}
\end{table}

\begin{table}
    \label{tab:ysi-data-distribution}
    \caption{Schema-Dependent X Schema-less Databases: Application Interface}
    \begin{center}
    \begin{tabular}{|p{210pt}|p{210pt}|}\hline
    Schema-Dependent Databases & Schema-less Databases\\\hline
    \begin{enumerate}
      \item Have their own specific API or through ODBC\footnote{Open Database
      Connectivity};
      \item Data is stored in a format that represents its natural structure,
      and for this reason, in a single or distributed fashion.
    \end{enumerate} 
    & 
    \begin{enumerate}
      \item Systems tend to provide SOAP/REST \cite{http-rest};
      \item \cite{db-is-rdbs-dommed} claims that data is stored in a more
      effective way, requiring only code plumbing for the relational code;
    \end{enumerate}
    \\\hline
    \end{tabular}
    \end{center}
\end{table}

@@

\begin{itemize}
  \item \textbf{MySQL}: Supports
  \item \textbf{Microsoft SQL}: Supports
  \item \textbf{Berkeley TinyDB}: NO Supports
  \item \textbf{MongoDB}: Supports
  \item \textbf{CouchDB}: Supports
  \item \textbf{IBM DB2}: Supports
\end{itemize}
